### 2022-02-14 introduction/introduction.tex
more

Proof assistants (or interactive theorem provers) are programs which allow users
to describe a theory formally and to prove the statements about this theory
using rules of the inference system the proof assistant is based on. These
programs can then check if the given proof is correct. As their core is usually
small and well-tested, proofs that pass the checking phase are considered
trustworthy.

They proved indispensible in proving very complex mathematical theorems in a
trustworthy manner. For example, four-color theorem \cite{gg-2008-four-color}
and Kepler conjecture \cite{th-2015-kepler} were proven inside proof assistants,
ending the need to trust hundreds of pages of hard mathematical proofs. Proof
assistants are also used to show correctness of complex software such as
compilers \cite{xl-09-compcert} or operating systems kernels \cite{gk-09-sel4}.
This verified software can be used in safety-critical scenarios such as nuclear
plants or autonomous vehicles.

@@

>>>>

Proof assistants (or interactive theorem provers) are programs which allow users
to describe a theory formally and to write proofs of the statements about this theory
using inference system rules of the proof assistant. These
programs can then check if the given proof is correct. As their core is usually
small and well-tested, proofs that pass the checking phase are considered
trustworthy.

Proof assistants showed indispensible in proving very complex mathematical theorems in a
trustworthy manner. For example, four-color theorem \cite{gg-2008-four-color}
and Kepler conjecture \cite{th-2015-kepler} were proven inside proof assistants,
ending the need to trust hundreds of pages of hard mathematical proofs. Proof
assistants are also used to show correctness of complex software such as
compilers \cite{xl-09-compcert} or operating systems kernels \cite{gk-09-sel4}.
This verified software can be used in safety-critical scenarios such as nuclear
plants or autonomous vehicles.

@@

This approach proved useful: around 69\% of the simple proof goals can be
discharged automatically using Sledgehammer \cite{kb-13-mash}. However, there is
clearly some untapped potential in the approach as it communicates with
automatic theorem provers through a translation. To bridge this translation gap,
attempts were made to use higher-order automatic theorem provers instead of
first-order ones \cite{ns-13-leo2sh}. This proved unfruitful as state-of-the-art
higher-order of the time theorem provers were primarily designed for small and tricky
problems. On the other hand, problems coming from hammers consist of mostly first-order formulas, are rather large,
and mostly require simple higher-order proof steps.

Thus, from a standpoint of proof assistant ideal automatic theorem prover fulfills
the following wishlist:
\begin{itemize}
  \item Performs as best first-order provers
  \item Supports higher-order logic
  \item Works well on large problems
\end{itemize}
This brings us to the topic of this thesis: \emph{How can we develop a theorem provers that fulfills this wishlist?}
